Quantifying social distancing behaviour across Australia in response to COVID-19

### 2020-04-30 manuscript/ms.Rmd

All of these datastreams are derived from apps that are primarily smartphone-based. As such they overrepresent demographics that are more likely to own smartphones and have account with these technology companies. Data on the demographics of the users that contributed to these particular datastreams are not available. However rough public estimates suggest that 90% of Facebook users are aged between 18 and 64. Younger and older demographics are underrepresented, with 4.5% facebook users vs 16% in the general population over the age of 65, and 5.6% vs 20% under the age of 18. Given that school-age children are underrepresented in COVID-19 cases in Australia, and that those over 65 are likely to be less mobile, these data may be unintentionally biased towards those contributing most strongly to transmission.
Apple's and Citymapper's direction requests likely represent deliberate visits to a specific location, and a location that the user either does not usually visit (and therefore requires directions) or would benefit from live traffic updates for.
Whilst Google's index of the time spent in residential places is likely to be a good proxy for people staying at home, it does not exclude the possibility that this time is spent in another person's household.

@@

The datastreams provided by Apple, Citymapper and Google are all publicly available for the express purpose of supporting public heath bodies in their response to COVID-19. The datastreams provided by Facebook are provided to researchers via Facebook's Data for Good  programme. All of these datasets are fully anonymised and aggregated at the level of either states and territories or Australia's 4 largest cities, over each day. The large-scale aggregation of these datasets ensures the privacy of users - the smallest population is that of the Northern Territory: over 245,000.

>>>>

All but one of these datastreams are derived from primarily smartphone-based apps. As such, they overrepresent demographics that are more likely to own smartphones and have accounts with these technology companies. In Australia, the penetration of smartphones in the population is close to saturation. Detailed data on the demographics of the users that contributed to these particular datastreams are not available for Australia. However, rough global estimates suggest that 90% of Facebook users are aged between 18 and 64. Younger and older demographics are underrepresented, with 4.5% Facebook users vs 16% in the general population over the age of 65, and 5.6% vs 20% under the age of 18. Given that school-age children are underrepresented in COVID-19 cases in Australia, and that those over 65 are likely to be less mobile, these data may be unintentionally biased towards those contributing most strongly to transmission.
Deatils about the nature of data collection and processing of the Apple's and Citymapper's direction requests are also unclear. They likely represent deliberate visits to a specific location, and a location that the user either does not usually visit (and therefore requires directions) or forwhich the user would would benefit from live traffic updates.
Google - through its extensive user base that includes all Android OS phoine users is likely to capture well information across demographics. Whilst Google's index of the time spent in residential places is therefore likely to be a good proxy for the amount of people staying at home in the general population, it does not exclude the possibility that this time is spent in another person's household. Such effects can, in general, not be well filtered out.

Finally, the data provided by NBN are the most distinct from the rest of the data analysed here -- they are guaranteed to be generated by users taht are static at home. They only include domestic users of NBN connections (although a number of small numer of self-emplyed customers may profit from domestic pricing plans). While this dataset is unable to provide an indication of mobility, it serves as a good proxy for a change in internet access from home, which is hypothesized to directly correlate with active internet use. We assume here that the size of a standard accessed item ( website, movie, music file) has not increased substantially in the monitored period, and therefore any change in data volume transferred relates direclty to active interaction with data.

@@

The datastreams provided by Apple, Citymapper and Google are all publicly available for the express purpose of supporting public heath bodies in their response to COVID-19. The datastreams provided by Facebook are provided to researchers via Facebook's Data for Good  programme. All of these datasets are fully anonymised and aggregated at the level of either states and territories or Australia's 4 largest cities, over each day. The large-scale aggregation of these datasets ensures the privacy of users - the smallest population is that of the Northern Territory: over 245,000.
The NBN dataset is provided exclusively to the University of Melbourne research team (contact M. Tomko).
